---
title: "2008-2013 Undiagnosed and Incidence for Philadelphia"
author: "Jeanette Birnbaum"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
      toc: true
      toc_depth: 2
      number_sections: true

---

```{r edit, include=FALSE, echo=FALSE}
#############################################################
# NOTES
#############################################################
# As of 11/4/16, the way use this file is to set runNew to TRUE
# to get fresh results. runNew to FALSE will used saved results

#############################################################
# EDIT THESE PARAMETERS
#############################################################
# Run a fresh estimation? And, define named vector of cases
runNew <- TRUE
estimation_cases <- c(`Base Case`='base_case', `Upper Bound`='upper_bound')

# Specify year min and max for analysis
# These have been determined from philly_1.pdf, the descriptives report
year_min <- 2008
year_max <- 2013

# Specify time step
diagInterval <- 0.25

# Local path of the Undiagnosed repo (that contains "wa", "templates", etc)
undx_repo <- '/Users/jeanette/Dropbox/School/PhD/HIV_WA'

# Within above repo, path to data files
data_path <- 'data/philly_ehars_formatted.csv'
prev_path <- 'data/philly_plwh.csv'
results_path <- 'philly/incidence_2008-2013/philly_2008-2013.csv'

```

```{r setup, include=FALSE}
#############################################################
# TECHNICAL SETUP
#############################################################

# Load libraries
# Install Rjkb using 
#   devtools::install_github('netterie/Rjkb/Rjkb')
library(Rjkb)
library(knitr)
library(gridExtra)
library(kableExtra)

# Read chunks
# read_chunk(file.path(undx_repo, run_path))

# Knitr options
knitr::opts_chunk$set(echo = FALSE)
options(knitr.table.format = "latex") 
```

```{r data, include=FALSE}
#############################################################
# AUTOMATED DATA LOADING AND FILE PATH SETUP
#############################################################

# Install the package
# devtools::install_github('hivbackcalc/package1.0/HIVBackCalc')
library(HIVBackCalc)

# Load the data and format it; load the HIVBackCalc package.
# Update the package before running this script - don't source interim work.
setup_hivbackcalc(workd=undx_repo,
                  datafile=data_path,
                  source_these='',
                  load_package=TRUE)

# Apply year min and max - why wasn't this code in this file before 10/26/18? 
dataf <- subset(dataf, yearDx>=year_min & yearDx<=year_max)

# Apply mode restrictions
dataf <- subset(dataf, new_mode!='Adult IDU' & 
  new_mode!='Perinatal exposure, HIV diagnosed at age 13 years or older')

# Make some unexpected adjustments to variable formats...
dataf <- transform(dataf,
                   everHadNegTest=as.logical(everHadNegTest),
                   infPeriod=as.numeric(infPeriod))


# Save a copy of full dataset and original results path
datafOrig <- dataf
results_path_orig <- results_path
prev_path_orig <- prev_path

# Read in the prevalence data - note the importance of check.names
prev_data <- read.csv(file.path(undx_repo, prev_path), 
                      check.names=FALSE,
                      header=TRUE)
```

```{r run, include=FALSE}
all <- runAnalysis(dataf, c(`Mode`='mode2'),
                   subvar='mode2', intLength=diagInterval, 
                   cases=estimation_cases,
                   runEstimation=runNew,
                   savedEstimation=!runNew,
                   prev=prev_data,
                   save=gsub('.csv', '_allKC.csv', 
                              file.path(undx_repo, results_path)))


# Compile and clean results
prettyprev <- all$resultsCompiled$trueprev

# Add PLWHA to the Base Case. Then to the Upper Bound in a 2nd version.
prettyprev <- within(prettyprev, {
                         `Diagnoses/Case`=as.character(`Diagnoses/Case`)
                         `Diagnoses/Case`=ifelse(`Diagnoses/Case`=='PLWHA', 'Base Case',
                                                 `Diagnoses/Case`)
                         
                   })

prettyprevUB <- all$resultsCompiled$trueprev
prettyprevUB <- within(prettyprevUB, {
                         `Diagnoses/Case`=as.character(`Diagnoses/Case`)
                         `Diagnoses/Case`=ifelse(`Diagnoses/Case`=='PLWHA', 
                                                 'Upper Bound',
                                                 `Diagnoses/Case`)
                         
                   })

# Melt and reshape?
longall <- subset(melt(prettyprev, id.vars=c('Subgroup', 'Diagnoses/Case', 
                                          'Estimate', 'Year')), 
               variable=='Mean')
longallUB <- subset(melt(prettyprevUB, id.vars=c('Subgroup', 'Diagnoses/Case', 
                                          'Estimate', 'Year')), 
               variable=='Mean')

wideall <- dcast(longall, Subgroup+Year+`Diagnoses/Case`~Estimate, value.var='value')
sublevels=c('MSM', 'non-MSM', 'Total-stratified')
sublabels=c('MSM', 'non-MSM', 'Total')
wideall$Subgroup <- factor(wideall$Subgroup, levels=sublevels, labels=sublabels)
wideall <- arrange(wideall, Year, Subgroup)

wideallUB <- dcast(longallUB, Subgroup+Year+`Diagnoses/Case`~Estimate, value.var='value')
wideallUB$Subgroup <- factor(wideallUB$Subgroup, levels=sublevels, labels=sublabels)
wideallUB <- arrange(wideallUB, Year, Subgroup)

write.csv(wideall, file.path(undx_repo, 
                    gsub('.csv', '_final.csv', results_path)),
                    row.names=FALSE,
                    na='')
```

```{r note1}
# Note that \newpage works because we're knitting to pdf. See 
# http://stackoverflow.com/questions/25240541/how-to-add-newpage-in-rmarkdown-in-a-smart-way
```

\newpage


# Diagnoses
## Analytic sample
Analytic data set has `r nrow(dataf)` cases for `r year_min`-`r year_max`. 

## Diagnoses over time by MSM subgroups
```{r dx1,  plot=TRUE}
all$dx$plot + aes(color=mode2)
```

# Subgroup Sizes and Testing Histories

```{r th1, results='asis'}
kable(all$th$table[,2:ncol(all$th$table)], caption='Column Percent shows the composition of the total sample. The Percent Yes, Percent No and Percent Missing columns indicate the row percents of the three possible testing history statuses within each subgroup', booktabs=T)
```


# Time from Infection to Diagnosis (TID)
The plots below show the TID curves for the MSM subgroups. BC=Base Case and UB=Upper Bound. For each subgroup, the upper plot panel is the probability of diagnosis curve (the pdf) and the lower plot panel is the survivor curve (1-cdf). The lower plot panel's y-axis thus indicates the fraction remaining undiagnosed at a given number of years since infection (x-axis).

```{r tid1, plot=TRUE, fig.height=6}
groups <- names(all$tid$separate)
grid.arrange(
                  plot(all$tid$separate[[1]],diagInterval, cases=c('BC', 'UB')) + 
                      ggtitle(groups[1]),
                  plot(all$tid$separate[[2]],diagInterval, cases=c('BC', 'UB')) +
                      ggtitle(groups[2]),
             ncol=2)
```


# Incidence and undiagnosed counts
The upper plot panels show diagnoses and estimated incidence, while the lower plot panels show undiagnosed cases. The upper panels have different y-scales.

```{r inc1, plot=TRUE, fig.height=6}
all$resultsPlots[['Total-stratified']]
```


# Undiagnosed fractions

10/26/18 WARNING: PLWH DATA ARE FAKE - I don't have real data for Philly

These plots show the total number of PLWH (x-axis) broken down by diagnosed and undiagnosed (colors) for the Base Case versus Upper Bound estimates (panels). Undiagnosed fractions are indicated by the percent labels. 

```{r frac1, plot=TRUE, fig.height=3.75}
all$resultsPrevPlots[['Total-stratified']]
```


