---
title: "2016 Incidence Estimates for King County"
author: "Jeanette Birnbaum"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
      toc: true
      toc_depth: 2
      number_sections: true

---

```{r edit, include=FALSE, echo=FALSE}
rm(list=ls())
#############################################################
# NOTES
#############################################################
# 2/21/18 - Using this script to extract incidence for Darcy
# 
# As of 11/4/16, the way use this file is to set runNew to TRUE
# to get fresh results. runNew to FALSE will used saved results

#############################################################
# EDIT THESE PARAMETERS
#############################################################
# Run a fresh estimation? And, define named vector of cases
runNew <- TRUE
estimation_cases <- c(`Base Case`='base_case', `Upper Bound`='upper_bound')

# Specify year min and max for analysis
year_min <- 2011
year_max <- 2016

# Specify time step
diagInterval <- 0.25

# Local path of the Undiagnosed repo (that contains "wa", "templates", etc)
undx_repo <- '/Users/jeanette/Dropbox/School/PhD/HIV_WA'

# Within above repo, path to data files
data_path <- 'data/kc_2017_jbexport.001to.04_formattedByApp.csv'
prev_path <- 'data/kc_2017_PLWHA.csv'
results_path <- 'kc/undx2016/results/undx2016.csv'

```

```{r setup, include=FALSE}
#############################################################
# TECHNICAL SETUP
#############################################################

# Load libraries
# Install Rjkb using 
#   devtools::install_github('netterie/Rjkb/Rjkb')
library(Rjkb)
library(knitr)
library(gridExtra)
library(kableExtra)

# Read chunks
# read_chunk(file.path(undx_repo, run_path))

# Knitr options
knitr::opts_chunk$set(echo = FALSE)
options(knitr.table.format = "latex") 
```

```{r data, include=FALSE}
#############################################################
# AUTOMATED DATA LOADING AND FILE PATH SETUP
#############################################################

# Install the package
# devtools::install_github('hivbackcalc/package1.0/HIVBackCalc')

# Load the data and format it; load the HIVBackCalc package.
# Update the package before running this script - don't source interim work.
setup_hivbackcalc(workd=undx_repo,
                  datafile=data_path,
                  source_these='',
                  load_package=TRUE)

# Make some unexpected adjustments to variable formats...
dataf <- transform(dataf,
                   everHadNegTest=as.logical(everHadNegTest),
                   infPeriod=as.numeric(infPeriod))


# Save a copy of full dataset and original results path
datafOrig <- dataf
results_path_orig <- results_path
prev_path_orig <- prev_path

# Read in the prevalence data - note the importance of check.names
prev_data <- read.csv(file.path(undx_repo, prev_path), 
                      check.names=FALSE,
                      header=TRUE)
```

```{r run, include=FALSE}
all <- runAnalysis(dataf, c(`Mode`='mode2'),
                   subvar='mode2', intLength=diagInterval, 
                   cases=estimation_cases,
                   runEstimation=runNew,
                   savedEstimation=!runNew,
                   prev=prev_data,
                   save=NULL)


# Incidence and undiagnosed counts
Save results that include incidence

```{r inc1, plot=TRUE, fig.height=6}
inc_path <- gsub('2016.csv', '2016_incidence.csv', results_path_orig)
tosave <- all$resultsCompiled$results$resultsAll
colnames(tosave) <- c('Quarter-Year', 'Measure', 'Detail', 'non-MSM', 'MSM', 'Total')
tosave[,4:6] <- round(tosave[,4:6], 1)

tosave <- subset(tosave, Measure=='Diagnoses and Incidence', select=-Measure)
tosave <- transform(tosave, Detail=gsub('Base Case', 'Base Case Incidence', Detail))
tosave <- transform(tosave, Detail=gsub('Upper Bound', 'Upper Bound Incidence', Detail))

colnames(tosave) <- c('Quarter-Year', 'Measure', 'non-MSM', 'MSM', 'Total')
write.csv(tosave, file=file.path(undx_repo, inc_path), row.names=FALSE)
```


