---
title: "Format WA Data - 2018 Estimates"
author: "Jeanette Birnbaum"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output:
  pdf_document:
      toc: true
      toc_depth: 2
      number_sections: true
---

```{r edit, include=FALSE, echo=FALSE}
rm(list=ls())
#############################################################
# EDIT THESE PARAMETERS
#############################################################
# Specify year min and max for analysis
year_min <- 2005
year_max <- 2018

# Local path of the Undiagnosed repo (that contains "wa", "templates", etc)
undx_repo <- '/Users/jeanette/Dropbox/School/PhD/HIV_WA'

# Within above repo, path to data file and formatting file
data_path <- 'data/wa_backcalc_data_201908.csv'
format_path <- 'wa/code/formatWA.R'
```

```{r setup, include=FALSE, echo=FALSE}
# Empty workspace and load libraries
# Install Rjkb using 
#   devtools::install_github('netterie/Rjkb/Rjkb')
library(Rjkb)
library(knitr)

# Knitr options
knitr::opts_chunk$set(echo = TRUE)

# Formatting code
read_chunk(file.path(undx_repo, format_path))
printSummaries <- TRUE
```

```{r data, include=FALSE, echo=FALSE}
# Load the data and format it; load the HIVBackCalc package.
# Update the package before running this script - don't source interim work.
setup_hivbackcalc(workd=undx_repo,
                  datafile=data_path,
                  source_these='',
                  load_package=TRUE)
```

# Raw Data Overview

## Sample Size
```{r overview, echo=FALSE}
```

N = `r Nobs0`

## Variable list
```{r rprintSummaries, echo=FALSE, include=FALSE}
```

```{r str}
str(dataf)
```
## Variable summaries
```{r summarize, echo=FALSE}
```
# Subset based on hst=WA and year

## First, split the combined year-quarter of diagnosis and AIDS variables
```{r split_yr_qtr, include=TRUE}
```

## Subset the data based on hst=WA and year
```{r subset_initial}
```
Excluded `r Nobs0-Nobs1` cases based on year and hst restrictions and missingness in age and year of diagnosis.

## New sample size
New sample size is `r Nobs1`

# Year and quarter of diagnosis: cleaning it up
## Years represented
```{r yearDx, echo=FALSE}
table(dataf$yearDx, useNA='ifany')
```

## Quarters represented
```{r quarterDx, echo=FALSE}
table(dataf$quarterDx, useNA='ifany')
```

## Distribute unknown quarters uniformly across Q1-Q4
```{r impute_qtr}
```

# Tabulate and collapse race and mode of diagnosis variables
## Race and mode by year
```{r investigate_race_mode}
    table(dataf$new_race, dataf$yearDx, useNA='ifany')
    table(dataf$new_mode, dataf$yearDx, useNA='ifany')
```
## Collapse
```{r collapse}
```

```{r investigate_race_mode_again}
    table(dataf$race, dataf$yearDx, useNA='ifany')
    table(dataf$mode, dataf$yearDx, useNA='ifany')
    table(dataf$mode2, dataf$yearDx, useNA='ifany')
```

# AIDS at Diagnosis

## AIDS at initial diagnosis?
```{r aidsAtDx, echo=FALSE}
    table(dataf$aidsAtDx, useNA='ifany')
```

## Years of AIDS diagnosis represented:
```{r yearAids, echo=FALSE}
    table(dataf$yearAids, useNA='ifany')
```

## Quarters of AIDS diagnosis represented:
```{r quarterAids, echo=FALSE}
    table(dataf$quarterAids, useNA='ifany')
```

# Ever had a last negative test (everHadNegTest)

## Coding
This variable will be coded as Yes=TRUE, No=FALSE, and Don't Know/Refused/Missing=NA

```{r everHadNegTest}
```

## Make compatible with recorded LNT dates

### Change incorrect FALSEs
We have `r checkEver[1,1]` cases with everHadNegTest=FALSE and `r checkEver[3,1]` with everHadNegTest=NA but have a time since last negative test. Change their everHadNegTest flag.

```{r fix_everHadNegTest_toTRUE}
```
### Change incorrect TRUEs
We have `r checkEver[2,2]` cases who have everHadNegTest=TRUE but have NO time since last negative test. Change their everHadNegTest flag. Change, 9/27/17 - previously was setting to false; now, set to NA.

```{r fix_everHadNegTest_toNA}
```

### Check

```{r check_everHadNegTest}
```

# Time since last negative test (infPeriod)

## Apply age-16 assumption and summarize
```{r infPeriod}
```

```{r summaryInf}
summary(dataf$infPeriod,digits=3)
```

## Diagnoses younger than 16

```{r infPeriod_investigate_neg}
```

There are `r sum(a1['TRUE',c(1,3)])` cases who do not have a date of last negative test and may not fit the assumption of TID=age-16. Of those, `r sum(a1['TRUE',c(1,3)])-sum(a2['TRUE',c(1,3)])` are age 16 at diagnosis and will have TID=0 using this assumption. Primary mode of transmission is Ped ('Perinatal or pediatric').

```{r infPeriod_cap_neg}
```
Excluded `r Nobs1-Nobs2` cases due to age $\leq$ 16 and no observed infPeriod data.

## Maximum window of 18 years
```{r infPeriod_cap_tested}
```
Among those with everHadNegTest=TRUE, we capped `r check_cap1[2,1]` cases at aidsUB.

```{r infPeriod_cap_nottested}
```
Among those with everHadNegTest=FALSE, no one had an original TID value.

```{r infPeriod_cap_NAtested}
```
Among those with everHadNegTest=NA, no one had an original TID value.

# Final analytic dataset
## Reminder of data cleaning
Final subset is of size `r nrow(dataf)`
* Diagnoses included: 
    - Year: non-missing, and 2005 onwards
    - Occurred in WA state
    - Excluded `r Nobs0-Nobs1` cases based on year and hst restrictions (no missingness in age and year of diagnosis in data for 2015 estimates).
* Ages included:
    - If missing age, must have recorded time of last negative test
    - If age $\leq$ 16, must have recorded time of last negative test
    - Excluded `r Nobs1-Nobs2` cases due to age $\leq$ 16 and no observed LNT. 

## Variable summaries
```{r final_summary, echo=FALSE}
```

## Cross-tabulation of KC diagnosis counts for SHAMP
```{r shamp, echo=TRUE}
# See diagnoses by year
table(dataf$yearDx)

# Subset to non-MSM in KC and look again
# Note that "Hetero" excludes "Blood/Needle"
nonMSMkc <- subset(dataf, dx_in_king=='Y' & mode2=='non-MSM')
table(nonMSMkc$yearDx)

# Create race3 which is just White/Black/Other
nonMSMkc$race3 <- ifelse(nonMSMkc$race!='White' & nonMSMkc$race!='Black', 'Other', as.character(nonMSMkc$race))
nonMSMkc$race3 <- factor(nonMSMkc$race3, levels=c('White', 'Black', 'Other'),
                         labels=c('White', 'Black', 'Other'))
table(nonMSMkc$race3, nonMSMkc$yearDx)
table(nonMSMkc$race3, nonMSMkc$timeDx)
```
```{r shamp-plot}
ggplot(nonMSMkc, aes(x=timeDx, fill=birth_sex)) + 
    geom_histogram(position='dodge', binwidth=0.25) + 
    facet_grid(race3~.) +
    scale_x_continuous(breaks=seq(2005,2018,by=1)) +
    theme(axis.text.x=element_text(angle=90,hjust=1)) +
ggtitle('Quarterly non-MSM HIV diagnoses in KC, by race and sex')
```

```{r shamp-plot2}
ggplot(nonMSMkc, aes(x=timeDx, fill=birth_sex)) + 
    geom_histogram(position='dodge', binwidth=0.5) + 
    facet_grid(race3~.) +
    scale_x_continuous(breaks=seq(2005,2018,by=1)) +
    theme(axis.text.x=element_text(angle=90,hjust=1)) +
ggtitle('Half-year non-MSM HIV diagnoses in KC, by race and sex')
```

```{r shamp-plot3}
ggplot(nonMSMkc, aes(x=timeDx, fill=birth_sex)) + 
    geom_histogram(position='dodge', binwidth=1) + 
    facet_grid(race3~.) +
    scale_x_continuous(breaks=seq(2005,2018,by=1)) +
    theme(axis.text.x=element_text(angle=90,hjust=1)) +
ggtitle('Yearly non-MSM HIV diagnoses in KC, by race and sex')
```