---
title: Integrating CD4 data into undiagnosed estimates
author: "Jeanette Birnbaum"
date: "October 3, 2016"
output: pdf_document
---


```{r setup, include=FALSE, echo=FALSE, cache=FALSE}
# Reminder: slide types are alert, prompt, section and sub-section
#############################################################
# SETUP
#############################################################
rm(list=ls())
# TEMPORARY: SOURCE FUNCTIONS
source('/Users/jeanette/Dropbox/School/PhD/HIV_WA/HIVBackCalc/R/internal_fxns.R')

# Change year min and max
year_min <- 2005
year_max <- 2014

# Load libraries, data and data-cleaning file
# Eventually this function should return the cleaned dataset,
# but data-cleaning has the name hardcoded as msm and I'm not
# going to generalize that right now
setup_hivbackcalc(workd='/Users/jeanette/Dropbox/School/PhD/HIV_WA',
                  datafile='data/wa_backcalc_data_201602.csv',
                  source_these='analysis_WA/format_data.R',
                  load_package=TRUE,
                  packagefile='HIVBackCalc/R/internal_fxns.R')

    library(xtable)
    library(gridExtra)
    library(plyr)
    library(reshape2)
    library(ggplot2)
    library(vcd)


#############################################################
# KNITR
#############################################################

library(knitr)
knit_hooks$set(inline = function(x) {
                    prettyNum(round(x,2), big.mark=",")
                  })
# set global chunk options
opts_chunk$set(fig.path='figure/minimal-', fig.align='center', 
               fig.show='hold', concordance=TRUE, external=TRUE,
               tidy=TRUE, size='footnotesize', verbose=TRUE)
options(formatR.arrow=TRUE,width=80,digits=7)

read_chunk(file.path(workd,'analysis_WA/2016_CD4Case.R'))
```

```{r cd4meds, include=FALSE, echo=FALSE}
```

```{r transformations, include=TRUE, echo=FALSE}
# Note that this MUST come after cd4meds, because dataf is 
# redefined more completely. cd4meds is preserved as-is so 
# I don't break the CD4Case report (v2 I believe)
```


## Project Goal

**To use CD4 data to _increase the precision_ of undiagnosed estimates from the testing history model** 

*Why* 
- Negative tests that are many years prior to diagnosis, i.e. "long infection windows", are not very informative regarding time of infection
- CD4 at diagnosis can indicate recency of infection

*How*
- From literature, identify typical times to infection for various CD4 counts
- Use this to update the probability of infection within long infection windows

*Impact in WA*
- Our pre-analysis of CD4 measurements in WA indicates that using CD4 will increase the precision of our undiagnosed estimates but will not significantly change the estimated number of undiagnosed cases
- The main impact of integrating CD4 will be to increase our confidence in the estimates 

## Reminder: the testing history method

## Base Case distributes probability uniformly across the infection window

_The hazard (instantaneous rate) of infection at any point within the window is 1/(window length), shown by the red line for two different window lengths. Time=0 refers to time of diagnosis._ 

```{r plotProbModel, include=TRUE, echo=FALSE}
```

```{r, include=TRUE, echo=FALSE, fig.width=8, fig.height=3.5, dpi=300}
par(mfrow=c(1,2))
plot.BC(2,xmax=4,ymax=2,t='Window=2 years')
plot.BC(4,xmax=4,ymax=2,t='Window=4 years')
```
The red line is lower when the window is longer, since the probability of infection is spread out over a longer time period.
